---
title: "Descriptive Analysis for Chicago Crime Data 2021"
author: "Anushka Jayasooriya"
date: "3/12/2022"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

Import the R libraries
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(scales)
```

Import the relevant datasets

```{r}
crime<-read.csv("./Data/2021_Crime_Data.csv")
crime_new<-read.csv("./Data/final_c.csv")
```

I have uploaded two datsets. crime is the pure dateset and crime_new is the extracted dataset using Python.
```{r}
head(crime)
```
```{r}
totNUm_1<-crime %>% summarise(count=n())
totNUm_1
```

The crime data set has included 205245 observations and 22 variables.
Now we will see the another set calles crime_new.
```{r}
head(crime_new)
```
It is a extracted dataset and it has contained some temporal variables.They are Month,Week,Day and Hour. Furthermore it has include Primary type.

Now we will see the relevant Primary type(Crime Species) counts on each wards
```{r}
crime_spe<-crime %>% 
  group_by(Primary.Type) %>% 
  summarise(Count = n())
```
```{r}
crime_spe$Primary.Type
```
According to the R code 31 Primary types are included.
```{r}
head(crime_spe)
```
```{r}
sum(crime_spe$Count)
```
Let we consier Primary type counts on each wards
```{r}
ward_counts<-crime %>% 
  group_by(Ward) %>% 
  summarise(Counts = n())
```

```{r}
head(ward_counts)
```
```{r}
ward_counts$Ward
```
There are 51 wards in this dataset and only one ward is unknown.

```{r}
sum(ward_counts$Counts)
```

Our main objective of this analysis is to get the total crime counts on each temporal variables for a particulat ward.
```{r}
crime_counts<-crime_new %>% 
  group_by(Ward,Month,Week,Day,Hour,Primary.Type) %>% 
  summarise(Counts = n())
```

```{r}
head(crime_counts)
```

```{r}
sum(crime_counts$Counts)
```
Let we see the most occured crimes in Chicago. Therefor we have to plot a bargraph on this senario.
```{r}
crime_spe %>% 
  ggplot(aes(x = reorder(Primary.Type ,Count), y = Count)) +
  geom_bar(stat = "identity", fill = "#759bb1") +
  labs(x ="Primary Type", y = "Number of Crimes", title = "") + 
  scale_y_continuous(label = comma) +
  coord_flip() 
```
According to this Bar plot the most crime is Theft. futhermore top five crimes are Theft,Battery,Criminal Damage,Assault and Deceptive Practice.

Now we will see how about the crime count variation among wards
```{r}
ward_counts %>% 
  ggplot(aes(x = reorder(Ward ,Counts), y = Counts)) +
  geom_bar(stat = "identity", fill = "#759bb1") +
  labs(x ="Primary Type", y = "Number of Crimes", title = "") + 
  scale_y_continuous(label = comma) +
  coord_flip() 
```
According to this bargraph most incidents occured in 28th ward and least one is 19th ward.

Let we see the temporal patterns using "crime counts" dataframe.

```{r}
head(crime_counts)
```
```{r}
crime_counts %>% 
  ggplot(aes(x = reorder(Month ,Counts), y = Counts)) +
  geom_bar(stat = "identity", fill = "#798cb1") +
  labs(x ="Month Number", y = "Total Crime Count", title = "Month Vs Count") + 
  scale_y_continuous(label = comma) +
  coord_flip() 
```
According that, October has occured more crime incidents and actually Autumn and Summar seasons.In February has occured less number of crimes.

```{r}
crime_counts %>% 
  ggplot(aes(x = reorder(Week ,Counts), y = Counts)) +
  geom_bar(stat = "identity", fill = "#798cb1") +
  labs(x ="Week Number", y = "Total Crime Count", title = "Week Vs Count") + 
  scale_y_continuous(label = comma) +
  coord_flip() 
```

Most incidents occure in 39th week in 2021 and 1st week has the lowest crime incidents.

```{r}
crime_counts %>% 
  ggplot(aes(x = reorder(Day ,Counts), y = Counts)) +
  geom_bar(stat = "identity", fill = "#798cb1") +
  labs(x ="Day of the Week", y = "Total Crime Count", title = "Day Vs Count") + 
  scale_y_continuous(label = comma) +
  coord_flip()
```

Friday and Saturday have occurred more crimes and Thursday has occured less incidents.

```{r}
crime_counts %>% 
  ggplot(aes(x = reorder(Hour ,Counts), y = Counts)) +
  geom_bar(stat = "identity", fill = "#798cb1") +
  labs(x ="Hour", y = "Total Crime Count", title = "Hour Vs Count") + 
  scale_y_continuous(label = comma) +
  coord_flip() 
```

In the mid night is the dangerous time and 5 is the lowest.We can say in the morning period has more crime cases and evening time slots have occurred less cases.




